---
title: "Salamander eye scaling and investment - loading data and initial graphs"
author: "Nicole Bednarik"
date: July 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: hide
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools) 
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
#library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
#library(evomap)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

## Morphological measurements

We recorded 6 morphological measurements from each salamander specimen: snout-vent length, wet mass, corneal diameters (left and right), and transverse eye diameters (left and right). Symmetrical measures (e.g. eye diameters) were averaged for each individual prior to further analysis. 

```{r load-data}

# Load raw data for measured frog specimens
salamanders_morph <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/salamanders_analysis.csv", header=TRUE, na.strings=c("", "NA", " ","<1"))) #load raw data

# Tidy dataset for analysis
salamanders.subset <- salamanders_morph %>%
  mutate(rootmass = Mass_g^(1/3)) %>% #adds cube root of mass
  mutate(eyemean = rowMeans(salamanders_morph[c('ED_right_mm', 'ED_left_mm')], na.rm=TRUE)) %>% #adds mean of L/R eyes
  mutate(cormean = rowMeans(salamanders_morph[c('CD_right_mm', 'CD_left_mm')], na.rm=TRUE)) %>%  #adds mean of L/R corneas
  mutate(eyemean = na_if(eyemean, "NaN")) %>% #remove missing values
  mutate(cormean = na_if(cormean, "NaN")) %>% #remove missing values
  select(Order, Suborder, Family, Subfamily, Genus, Species, Genus_Species, SVL_mm, Mass_g, rootmass, ED_right_mm, ED_left_mm, eyemean, CD_right_mm, CD_left_mm, cormean) #keeps only columns of interest for analyses
```

## Ecological traits

We coded 6 ecological traits for each species from published literature into binary categories, and then merged this data with our specimen measurements by genus and species. 

```{r trait-data, results = "hide"}

#Import ecological trait data
traits <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/salamanders_traits.csv"))

#Change "Unknown" trait categorizations to NA so they will be excluded from statistical comparisons
traits.subset <- traits %>%
  mutate(Aquatic = factor(na_if(Aquatic, "Unknown"))) %>%
  mutate(Semiaquatic = factor(na_if(Semiaquatic, "Unknown"))) %>%
  mutate(Terrestrial = factor(na_if(Terrestrial, "Unknown"))) %>%
  mutate(Scansorial = factor(na_if(Scansorial, "Unknown"))) %>%
  mutate(Fossorial = factor(na_if(Fossorial, "Unknown"))) %>%
  mutate(Subfossorial = factor(na_if(Subfossorial, "Unknown"))) %>%
  mutate(Cave_dwelling = factor(na_if(Cave_dwelling, "Unknown"))) %>%
  mutate(Nocturnal = factor(na_if(Nocturnal, "Unknown"))) %>%
  mutate(Diurnal = factor(na_if(Diurnal, "Unknown"))) %>%
  mutate(Paedomorphic = factor(na_if(Paedomorphic, "Unknown"))) %>%
  mutate(Metamorphosizing = factor(na_if(Metamorphosizing, "Unknown"))) %>%
  mutate(Direct_Development = factor(na_if(Direct_Development, "Unknown"))) %>%
  mutate(Free_living = factor(na_if(Free_living, "Unknown"))) %>%
  mutate(No_free_living = factor(na_if(No_free_living, "Unknown"))) %>%
  mutate(Lentic_water = factor(na_if(Lentic_water, "Unknown"))) %>%
  mutate(Lotic_water = factor(na_if(Lotic_water, "Unknown"))) %>%
  mutate(Both_water = factor(na_if(Both_water, "Unknown"))) %>%
  mutate(No_larva = factor(na_if(No_larva, "Unknown"))) %>%
  mutate(Females_larger = factor(na_if(Females_larger, "Unknown"))) %>%
  mutate(Males_larger = factor(na_if(Males_larger, "Unknown"))) %>%
  mutate(No_SSD = factor(na_if(No_SSD, "Unknown"))) %>%
  select(Genus_Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD)

# Merge morphology dataset with ecological trait dataset
salamanders <- merge(salamanders.subset, traits.subset, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)

# Check structure of final dataframe for analyses (numbers and factors in propoer columns)
str(salamanders)
```

## Sampling

We sampled all 9 extant caudata families, and morphological sampling for interspecific analysis ranged from n = 1 to n = 6 individuals for each species (based on availability in the collections).

```{r sampling, results = "hide"}

#Number of species and individuals sampled 
counts <-ddply(salamanders, .(salamanders$Family, salamanders$Subfamily, salamanders$Genus, salamanders$Species), nrow)
names(counts) <- c("Family","Subfamily", "Genus","Species","Sampled")

#create scrolling RMarkdown table of sampling
kable(counts[ , c("Family","Genus","Species","Sampled")], caption = "Species and sampling effort for morphological data from museum specimens.") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  scroll_box(height = "500px")

#export sampling data for supplement
write.csv(counts, file = "/Users/nicolebednarik/Desktop/salamander-eye-size-master/sampling.csv")

```

## Molecular phylogeny

Here we import an amphibian tree from Jetz and Pyron (2019) that has been pruned to match the salamander species in this dataset. 

```{r}
caudatatree <- read.nexus("/Users/nicolebednarik/Desktop/salamander-eye-size-master/caudata-tree")
plot.phylo(caudatatree, show.tip.label=TRUE)
```

# Interspecific allometry

## Species means for morphological comparisons

We then found species means for various morphological parameters. In order to keep as many specimens as possible in our comparisons, we created a new dataframe for each set of morphological comparisons, so that the few specimens with missing data for one parameter could still be used in comparisons not including that parameter. 

```{r species-av}
## SPECIES AVERAGES FOR VARIOUS COMPARISONS

##### eye diameter vs. snout vent length #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.edsvl <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(SVL_mm))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), svl_av = mean(SVL_mm), n = n())

## Merge data means with other column info
av.edsvl <- merge(av.edsvl, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edsvl <- av.edsvl %>% select(Genus_Species, eye_av, svl_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

#Load Adult Habitat dataset
adult_habitat <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/adult_habitat.csv"))

## Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.edsvl.habitat1 <- merge(av.edsvl, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.edsvl.habitat <- na.omit(av.edsvl.habitat1)

##Plot average eye diameter (mm) against SVL (mm)
ggplot(av.edsvl.habitat, aes(y=eye_av, x=svl_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=6, aes(color=Adult_habitat)) +
   theme_bw() +
  scale_color_manual(values=c("magenta", "7F00FF", "blue", "#E80A0A", "#00CC00")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "SVL (mm)") 
```

```{r}
##### eye diameter vs. the cube root of mass #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.edmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.edmass <- merge(av.edmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edmass <- av.edmass %>% select(Genus_Species, eye_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

# Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.edmass.habitat1 <- merge(av.edmass, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.edmass.habitat <- na.omit(av.edmass.habitat1)

#Plot average eye diameter (mm) against SVL (mm)
ggplot(av.edmass.habitat, aes(y=eye_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "Cube root of mass (g)") 
```

```{r}
##### cube root of mass vs. SVL #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.svlmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(SVL_mm) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(svl_av = mean(SVL_mm), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.svlmass <- merge(av.svlmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.svlmass <- av.svlmass %>% select(Genus_Species, svl_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD)

# Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.svlmass.habitat1 <- merge(av.svlmass, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.svlmass.habitat <- na.omit(av.svlmass.habitat1)
ggplot(av.svlmass.habitat, aes(y=svl_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Cube root of mass (g)") +
  scale_x_log10(name = "SVL (mm)") 
```

```{r}
##### cornea diameter vs. eye diameter #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.edcd <- salamanders %>% 
  filter(!(is.na(cormean)|is.na(eyemean))) %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!is.na(eyemean) | !is.na(cormean)) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), cornea_av = mean(cormean), n = n())

## Merge data means with other column info
av.edcd <- merge(av.edcd, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edcd <- av.edcd %>% select(Genus_Species, eye_av, cornea_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

# Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.edcd.habitat1 <- merge(av.edcd, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.edcd.habitat <- na.omit(av.edcd.habitat1)
ggplot(av.edcd.habitat, aes(y=cornea_av, x=eye_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Cornea diameter (mm)") +
  scale_x_log10(name = "Eye diameter (mm)") 
```


```{r}
##### cornea diameter vs. snout vent length #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.cdsvl <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(SVL_mm))) %>% 
  group_by(Genus_Species) %>%
  summarise(cornea_av = mean(cormean), svl_av = mean(SVL_mm), n = n())

## Merge data means with other column info
av.cdsvl <- merge(av.cdsvl, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.cdsvl <- av.cdsvl %>% select(Genus_Species, cornea_av, svl_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

# Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.cdsvl.habitat1 <- merge(av.cdsvl, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.cdsvl.habitat <- na.omit(av.cdsvl.habitat1)
ggplot(av.cdsvl.habitat, aes(y=cornea_av, x=svl_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Cornea diameter (mm)") +
  scale_x_log10(name = "SVL (mm)") 
```

```{r}
##### cornea diameter vs. cube root of mass #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.cdmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(cornea_av = mean(cormean), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.cdmass <- merge(av.cdmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.cdmass <- av.cdmass %>% select(Genus_Species, cornea_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD)

# Merge overall dataset with adult habitat information (and remove specimens with missing data)
av.cdmass.habitat1 <- merge(av.cdmass, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
av.cdmass.habitat <- na.omit(av.cdmass.habitat1)
ggplot(av.cdmass.habitat, aes(y=cornea_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Cornea diameter (mm)") +
  scale_x_log10(name = "Cube root of mass (g)") 
```


## Initial boxplot of eye diameter by adult habitat (not taking into account mass/SVL)

```{r}

ggplot(data = salamanders5, aes(x = Adult_habitat, y = eye_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun = mean, colour="black", geom="point", 
               shape=18, size=3, show.legend = FALSE) +
  geom_jitter(aes(color = Adult_habitat, text = Genus_Species), shape = 19, size = 3, alpha = 0.9, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_x_discrete(labels=c("Aq.", "Semiaq.", "Foss.", "Subfoss.", "Ground", "Scans.")) +
  ylab("Eye diameter (mm)") +
  xlab("Adult habitat") +
  theme(legend.position = "none")

```

