---
title: "Salamander eye scaling and investment"
author: "Nicole Bednarik"
date: July 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: hide
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools) 
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
#library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
#library(evomap)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

## Morphological measurements

We recorded 6 morphological measurements from each salamander specimen: snout-vent length, wet mass, corneal diameters (left and right), and transverse eye diameters (left and right). Symmetrical measures (e.g. eye diameters) were averaged for each individual prior to further analysis. 

```{r load-data}

# Load raw data for measured frog specimens
salamanders_morph <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/salamanders_analysis.csv", header=TRUE, na.strings=c("", "NA", " ","<1"))) #load raw data

# Tidy dataset for analysis
salamanders.subset <- salamanders_morph %>%
  mutate(rootmass = Mass_g^(1/3)) %>% #adds cube root of mass
  mutate(eyemean = rowMeans(salamanders_morph[c('ED_right_mm', 'ED_left_mm')], na.rm=TRUE)) %>% #adds mean of L/R eyes
  mutate(cormean = rowMeans(salamanders_morph[c('CD_right_mm', 'CD_left_mm')], na.rm=TRUE)) %>%  #adds mean of L/R corneas
  mutate(eyemean = na_if(eyemean, "NaN")) %>% #remove missing values
  mutate(cormean = na_if(cormean, "NaN")) %>% #remove missing values
  select(Order, Suborder, Family, Subfamily, Genus, Species, Genus_Species, SVL_mm, Mass_g, rootmass, ED_right_mm, ED_left_mm, eyemean, CD_right_mm, CD_left_mm, cormean) #keeps only columns of interest for analyses
```

## Ecological traits

We coded 6 ecological traits for each species from published literature into binary categories, and then merged this data with our specimen measurements by genus and species. 

```{r trait-data, results = "hide"}

#Import ecological trait data
traits <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/salamanders_traits.csv"))

#Change "Unknown" trait categorizations to NA so they will be excluded from statistical comparisons
traits.subset <- traits %>%
  mutate(Aquatic = factor(na_if(Aquatic, "Unknown"))) %>%
  mutate(Semiaquatic = factor(na_if(Semiaquatic, "Unknown"))) %>%
  mutate(Terrestrial = factor(na_if(Terrestrial, "Unknown"))) %>%
  mutate(Scansorial = factor(na_if(Scansorial, "Unknown"))) %>%
  mutate(Fossorial = factor(na_if(Fossorial, "Unknown"))) %>%
  mutate(Subfossorial = factor(na_if(Subfossorial, "Unknown"))) %>%
  mutate(Cave_dwelling = factor(na_if(Cave_dwelling, "Unknown"))) %>%
  mutate(Nocturnal = factor(na_if(Nocturnal, "Unknown"))) %>%
  mutate(Diurnal = factor(na_if(Diurnal, "Unknown"))) %>%
  mutate(Paedomorphic = factor(na_if(Paedomorphic, "Unknown"))) %>%
  mutate(Metamorphosizing = factor(na_if(Metamorphosizing, "Unknown"))) %>%
  mutate(Direct_Development = factor(na_if(Direct_Development, "Unknown"))) %>%
  mutate(Free_living = factor(na_if(Free_living, "Unknown"))) %>%
  mutate(No_free_living = factor(na_if(No_free_living, "Unknown"))) %>%
  mutate(Lentic_water = factor(na_if(Lentic_water, "Unknown"))) %>%
  mutate(Lotic_water = factor(na_if(Lotic_water, "Unknown"))) %>%
  mutate(Both_water = factor(na_if(Both_water, "Unknown"))) %>%
  mutate(No_larva = factor(na_if(No_larva, "Unknown"))) %>%
  mutate(Females_larger = factor(na_if(Females_larger, "Unknown"))) %>%
  mutate(Males_larger = factor(na_if(Males_larger, "Unknown"))) %>%
  mutate(No_SSD = factor(na_if(No_SSD, "Unknown"))) %>%
  select(Genus_Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD)

# Merge morphology dataset with ecological trait dataset
salamanders <- merge(salamanders.subset, traits.subset, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)

# Check structure of final dataframe for analyses (numbers and factors in propoer columns)
str(salamanders)
```

## Sampling

We sampled all 9 extant caudata families, and morphological sampling for interspecific analysis ranged from n = 1 to n = 6 individuals for each species (based on availability in the collections).

```{r sampling, results = "hide"}

#Number of species and individuals sampled 
counts <-ddply(salamanders, .(salamanders$Family, salamanders$Subfamily, salamanders$Genus, salamanders$Species), nrow)
names(counts) <- c("Family","Subfamily", "Genus","Species","Sampled")

#create scrolling RMarkdown table of sampling
kable(counts[ , c("Family","Genus","Species","Sampled")], caption = "Species and sampling effort for morphological data from museum specimens.") %>%
  kable_styling(full_width = F) %>%
  collapse_rows(columns = 1, valign = "top") %>%
  scroll_box(height = "500px")

#export sampling data for supplement
write.csv(counts, file = "/Users/nicolebednarik/Desktop/salamander-eye-size-master/sampling.csv")

```

## Molecular phylogeny

Here we import an amphibian tree from Jetz and Pyron (2019) that has been pruned to match the salamander species in this dataset. 

```{r}

caudatatree <- read.nexus("/Users/nicolebednarik/Desktop/salamander-eye-size-master/caudata-tree")
plot.phylo(caudatatree, show.tip.label=TRUE)
```

# Interspecific allometry

## Species means for morphological comparisons

We then found species means for various morphological parameters. In order to keep as many specimens as possible in our comparisons, we created a new dataframe for each set of morphological comparisons, so that the few specimens with missing data for one parameter could still be used in comparisons not including that parameter. 

```{r species-av}

## SPECIES AVERAGES (1-5 specimens per species) FOR VARIOUS COMPARISONS

##### eye diameter vs. snout vent length #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.edsvl <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(SVL_mm))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), svl_av = mean(SVL_mm), n = n())

## Merge data means with other column info
av.edsvl <- merge(av.edsvl, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edsvl <- av.edsvl %>% select(Genus_Species, eye_av, svl_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

##### eye diameter vs. the cube root of mass #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.edmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.edmass <- merge(av.edmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edmass <- av.edmass %>% select(Genus_Species, eye_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD) 

##### cube root of mass vs. SVL #####

# Take mean of replicate specimens for each species (and remove specimens with missing data for two traits of interest)
av.svlmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(SVL_mm) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(svl_av = mean(SVL_mm), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.svlmass <- merge(av.svlmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.svlmass <- av.svlmass %>% select(Genus_Species, svl_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, Both_water, No_larva, Females_larger, Males_larger, No_SSD)
```


```{r}
activity_period <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/activity_period.csv"))
salamanders2 <- merge(av.edmass, activity_period, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
salamanders3 <- na.omit(salamanders2)
ggplot(salamanders3, aes(y=eye_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Activity_period)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "Cube root of mass (g)") 

adult_habitat <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/adult_habitat.csv"))
salamanders4 <- merge(av.edmass, adult_habitat, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
salamanders5 <- na.omit(salamanders4)
ggplot(salamanders5, aes(y=eye_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Adult_habitat)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "Cube root of mass (g)") 

salamanders6 <- merge(av.edsvl, activity_period, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
salamanders7 <- na.omit(salamanders6)
ggplot(salamanders7, aes(y=eye_av, x=svl_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Activity_period)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "SVL (mm)") 

salamanders8 <- merge(av.svlmass, activity_period, by="Genus_Species", 
                   all.x = TRUE, all.y = FALSE)
salamanders9 <- na.omit(salamanders8)
ggplot(salamanders9, aes(y=rootmass_av, x=svl_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Activity_period)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Cube root of mass (g)") +
  scale_x_log10(name = "SVL (mm)") 

development <- data.frame(read.csv("/Users/nicolebednarik/Desktop/salamander-eye-size-master/development.csv"))
salamanders10 <- merge(av.edmass, development, by="Genus_Species", all.x = TRUE, all.y= FALSE)
salamanders11 <- na.omit(salamanders10)
ggplot(salamanders11, aes(y=eye_av, x=rootmass_av, text=Genus_Species)) +
  geom_point(alpha=0.9, size=4, aes(color=Development)) +
   theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  scale_y_log10(name = "Eye diameter (mm)") +
  scale_x_log10(name = "Cube root of mass (g)") 
```


```{r}

ggplot(data = salamanders5, aes(x = Adult_habitat, y = eye_av)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA) +
  stat_summary(fun = mean, colour="black", geom="point", 
               shape=18, size=3, show.legend = FALSE) +
  geom_jitter(aes(color = Adult_habitat, text = Genus_Species), shape = 19, size = 3, alpha = 0.9, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  theme(axis.text = element_text(angle = 0, size = 9)) +
  scale_x_discrete(labels=c("Aq.", "Semiaq.", "Foss.", "Subfoss.", "Ground", "Scans.")) +
  ylab("Eye diameter (mm)") +
  xlab("Adult habitat") +
  theme(legend.position = "none")

```

## Phylogenetic generalised least squares regressions

We fit PGLS models using the modified Feng et al. (2017) phylogeny and our averaged species data for morphological traits. 

### PGLS of log eye diameter vs. log cube root of mass

```{r pgls-edmass}

#fit PGLS model for log eye diameter vs. log cube root of mass using 'caper' package

#### check for best model of correlation structure ####

#Make list of taxa to drop (in tree but not in dataset)
drops <- setdiff(caudatatree$tip.label, as.character(av.edmass$mod_tiplabel))

#Drop unwanted tips from phylogeny
tree.edit <- drop.tip(phy = caudatatree, tip = drops) 

#Make row names of the species the phylogeny tip labels
rownames(av.edmass) <- av.edmass$mod_tiplabel

#Reorder species data to match phylogeny order
dat <-ReorderData(tree.edit, av.edmass, taxa.names="row.names")

#test whether BM or OU model is better fit for PGLS using the nlme package and AIC comparison

#Expected covariance under a pure Brownian model (Felsenstein 1985,	Martins and Hansen 1997)
pglsBM <- gls(log10(eye_av) ~ log10(rootmass_av), cor = corBrownian(phy = tree.edit), data = dat, method = "ML")

#Martins and Hansen's (1997) covariance structure (OU)
pglsOU<-gls(log10(eye_av) ~ log10(rootmass_av), cor = corMartins(1, phy = tree.edit, fixed = F), data = dat, method = "ML")

#Pagel's “lambda” Correlation Structure: The correlation structure from the present model is derived from the Brownian motion model by multiplying the off-diagonal elements (i.e., the covariances) by lambda. The variances are thus the same than for a Brownian motion model.
pglsPagel<-gls(log10(eye_av) ~ log10(rootmass_av), cor = corPagel(1, phy = tree.edit, fixed = F), data = dat, method = "ML")

anova(pglsBM, pglsOU, pglsPagel)

#### prep data ####

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.edmass, data.names = av.edmass$mod_tiplabel)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
edmass.comp <- comparative.data(phy = caudatatree, data = av.edmass, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
edmass.comp$dropped$tips #phylogeny
edmass.comp$dropped$unmatched.rows #dataset

#rescale tree to fit OU model with alpha = 0.5
frog.treeOU <- geiger::rescale(frog.tree, model = "OU", 0.5)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
edmass.compOU <- comparative.data(phy = frog.treeOU, data = av.edmass, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)


#for eye diameter vs. cube root of mass using the OU tree
pgls_edmassOU <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.compOU, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#### run PGLS model ####

#for eye diameter vs. cube root of mass using the Maximum Liklihood estimate of lambda
pgls_edmass <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#### check model assumptions ####

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass)
par(mfrow = c(1, 1))

#### model outputs ####

#print model output 
summary(pgls_edmass)

#quick plot of pgls
plot(log10(eye_av) ~ log10(rootmass_av), data = av.edmass)
abline(pgls_edmass)

#Likelihood plot for Pagel's lambda from the PGLS model of eye diameter vs. the cuberoot of mass. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.edmass <- pgls.profile(pgls_edmass, "lambda")
plot(lambda.edmass)

```

For the PGLS of log eye diameter vs. log cube root of mass, Pagel's lambda was `r pgls.confint(pgls_edmass, "lambda")$opt` with 95% confidence intervals of `r pgls.confint(pgls_edmass, "lambda")$ci.val`, indicating that there is high phylogenetic signal in the residuals of our PGLS model fit (which is our measure of relative eye investment).

The diagnostic plots for the model show that the data conformed to most model assumptions, though we do have one species with the absolute value of studentized residual exceeding 3, which could disproportionately influence the regression. To check whether this is happening, we re-ran the model iteratively with outliers removed to check the robustness of parameter estimates. 

```{r pgls-edmass-outliers, include = FALSE}

# refit pgls after removing extreme outlier/s observed in previous fit

#### Iteration 1 ####

#extract the phylogenetic residuals from the original model
phyres.edmass1 <- residuals(pgls_edmass, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass1 <- phyres.edmass1/sqrt(var(phyres.edmass1))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass1)<-rownames(pgls_edmass$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass1)[(abs(stphyres.edmass1)>3)]
 
#remove these two species from data/tree
edmass.comp_1 <- edmass.comp[-which(abs(stphyres.edmass1)>3), ]

#re-run pgls analysis (without the outliers)
pgls_edmass_1 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_1, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_1, main = "PGLS eye diameter v root mass (- outliers 1)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_1 , "lambda"))

#print model output 
summary(pgls_edmass_1)

#### Iteration 2 ####

#extract the phylogenetic residuals from the original model
phyres.edmass2 <- residuals(pgls_edmass_1, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass2 <- phyres.edmass2/sqrt(var(phyres.edmass2))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass2)<-rownames(pgls_edmass_1$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass2)[(abs(stphyres.edmass2)>3)]
 
#remove these two species from data/tree
edmass.comp_2 <- edmass.comp_1[-which(abs(stphyres.edmass2)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_2 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_2, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_2, main = "PGLS model for eye v root mass (- outliers 2)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_2 , "lambda"))

#print model output 
summary(pgls_edmass_2)

#### Iteration 3 ####

#extract the phylogenetic residuals from the original model
phyres.edmass3 <- residuals(pgls_edmass_2, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass3 <- phyres.edmass3/sqrt(var(phyres.edmass3))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass3)<-rownames(pgls_edmass_2$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass3)[(abs(stphyres.edmass3)>3)]
 
#remove these two species from data/tree
edmass.comp_3 <- edmass.comp_2[-which(abs(stphyres.edmass3)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_3 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_3, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_3, main = "PGLS model for eye v root mass (- outliers 3)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_3 , "lambda"))

#print model output 
summary(pgls_edmass_3)

#### Iteration 4 ####

#extract the phylogenetic residuals from the original model
phyres.edmass4 <- residuals(pgls_edmass_3, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass4 <- phyres.edmass4/sqrt(var(phyres.edmass4))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass4)<-rownames(pgls_edmass_3$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass4)[(abs(stphyres.edmass4)>3)]
 
#remove these two species from data/tree
edmass.comp_4 <- edmass.comp_3[-which(abs(stphyres.edmass4)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_4 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_4, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_4, main = "PGLS model for eye v root mass (- outliers 4)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_4 , "lambda"))

#print model output 
summary(pgls_edmass_4)

#### Iteration 5 ####

#extract the phylogenetic residuals from the original model
phyres.edmass5 <- residuals(pgls_edmass_4, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass5 <- phyres.edmass5/sqrt(var(phyres.edmass5))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass5)<-rownames(pgls_edmass_4$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass5)[(abs(stphyres.edmass5)>3)]
 
#remove these two species from data/tree
edmass.comp_5 <- edmass.comp_4[-which(abs(stphyres.edmass5)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_5 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_5, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_5, main = "PGLS model for eye v root mass (no outliers 5)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_5 , "lambda"))

#print model output 
summary(pgls_edmass_5)

#### Iteration 6 ####

#extract the phylogenetic residuals from the original model
phyres.edmass6 <- residuals(pgls_edmass_5, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass6 <- phyres.edmass6/sqrt(var(phyres.edmass6))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass6) <- rownames(pgls_edmass_5$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass6)[(abs(stphyres.edmass6)>3)]
 
#remove these two species from data/tree
edmass.comp_6 <- edmass.comp_5[-which(abs(stphyres.edmass6)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_6 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_6, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_6, main = "PGLS model for eye v root mass (no outliers)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_6 , "lambda"))

#print model output 
summary(pgls_edmass_6)

#### Iteration 7 ####

#extract the phylogenetic residuals from the original model
phyres.edmass7 <- residuals(pgls_edmass_6, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass7 <- phyres.edmass7/sqrt(var(phyres.edmass7))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass7)<-rownames(pgls_edmass_6$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass7)[(abs(stphyres.edmass7)>3)]

#remove these two species from data/tree
edmass.comp_7 <- edmass.comp_6[-which(abs(stphyres.edmass7)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_7 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_7, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_7, main = "PGLS model for eye v root mass (no outliers)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_7 , "lambda"))

#print model output 
summary(pgls_edmass_7)

#### Iteration 8 ####

#extract the phylogenetic residuals from the original model
phyres.edmass8 <- residuals(pgls_edmass_7, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass8 <- phyres.edmass8/sqrt(var(phyres.edmass8))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass8)<-rownames(pgls_edmass_7$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass8)[(abs(stphyres.edmass8)>3)]

#remove these two species from data/tree
edmass.comp_8 <- edmass.comp_7[-which(abs(stphyres.edmass8)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_8 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_8, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_8, main = "PGLS model for eye v root mass (no outliers)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_8 , "lambda"))
 
#print model output 
summary(pgls_edmass_8)

#### Iteration 9 ####

#extract the phylogenetic residuals from the original model
phyres.edmass9 <- residuals(pgls_edmass_8, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass9 <- phyres.edmass9/sqrt(var(phyres.edmass9))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass9)<-rownames(pgls_edmass_8$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass9)[(abs(stphyres.edmass9)>3)]

#remove these two species from data/tree
edmass.comp_9 <- edmass.comp_8[-which(abs(stphyres.edmass9)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_9 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_9, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_9, main = "PGLS model for eye v root mass (no outliers)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_9 , "lambda"))

#print model output 
summary(pgls_edmass_9)

#### Iteration 10 ####

#extract the phylogenetic residuals from the original model
phyres.edmass10 <- residuals(pgls_edmass_9, phylo = TRUE)

#standardize residuals by dividing by the square root of their variance
stphyres.edmass10 <- phyres.edmass10/sqrt(var(phyres.edmass10))[1] 

#match species names with standardized residuals
rownames(stphyres.edmass10)<-rownames(pgls_edmass_9$residuals)

#pull the names of the outlier species (where standardized res beyond ±3)
rownames(stphyres.edmass10)[(abs(stphyres.edmass10)>3)]

#remove these two species from data/tree
edmass.comp_10 <- edmass.comp_8[-which(abs(stphyres.edmass9)>3), ]

#re-run pgls analisys (without the outliers)
pgls_edmass_10 <- pgls(log10(eye_av) ~ log10(rootmass_av), 
               data = edmass.comp_10, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edmass_10, main = "PGLS model for eye v root mass (no outliers)")
par(mfrow = c(1, 1))

#lambda
plot(pgls.profile(pgls_edmass_10 , "lambda"))

#print model output 
summary(pgls_edmass_10)
```

As iterative removal of phylogenetic outliers has no major effect on model parameter estimates, we include all species in our PGLS of log eye diameter vs. log cube root of mass. 

### PGLS of log eye diameter vs. log snout-vent length 

```{r pgls-edsvl}

#fit PGLS model for log eye diameter vs. log SVL using 'caper' package

#### check for best model of correlation structure ####

#Make list of taxa to drop (in tree but not in dataset)
drops <- setdiff(frog.tree$tip.label, as.character(av.edsvl$mod_tiplabel))

#Drop unwanted tips from phylogeny
tree.edit <- drop.tip(phy = frog.tree, tip = drops) 

#Make row names of the species the phylogeny tip labels
rownames(av.edsvl) <- av.edsvl$mod_tiplabel

#Reorder species data to match phylogeny order
dat <-ReorderData(tree.edit, av.edsvl, taxa.names="row names")

#test whether BM or OU model is better fit for PGLS using the nlme package and AIC comparison

#Expected covariance under a pure Brownian model (Felsenstein 1985,	Martins and Hansen 1997)
pglsBM <- gls(log10(eye_av) ~ log10(svl_av), cor = corBrownian(phy = tree.edit), data = dat, method = "ML")

#Martins and Hansen's (1997) covariance structure (OU)
pglsOU<-gls(log10(eye_av) ~ log10(svl_av), cor = corMartins(1, phy = tree.edit, fixed = F), data = dat, method = "ML")

#Pagel's “lambda” Correlation Structure: The correlation structure from the present model is derived from the Brownian motion model by multiplying the off-diagonal elements (i.e., the covariances) by lambda. The variances are thus the same than for a Brownian motion model.
pglsPagel<-gls(log10(eye_av) ~ log10(svl_av), cor = corPagel(1, phy = tree.edit, fixed = F), data = dat, method = "ML")

anova(pglsBM, pglsOU, pglsPagel)

#### prep data ####

#check that names match in dataframe and tree
name.check(phy = frog.tree, data = av.edsvl, data.names = av.edsvl$mod_tiplabel)
  
#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
edsvl.comp <- comparative.data(phy = frog.tree, data = av.edsvl, 
                            names.col = mod_tiplabel, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
edsvl.comp$dropped$tips #phylogeny
edsvl.comp$dropped$unmatched.rows #dataset

#### run PGLS model ####

#for log eye diameter vs. log svl using the Maximum Liklihood estimate of lambda
pgls_edsvl <- pgls(log10(eye_av) ~ log10(svl_av), 
               data = edsvl.comp, 
               lambda = "ML", #uses Maximum Liklihood estimate of lambda
               param.CI = 0.95)

#### check model assumptions ####

#diagnostic plots
par(mar = c(4,4,2,2))
par(mfrow = c(2, 2))
plot(pgls_edsvl, main = "PGLS model for mean eye diameter vs. svl")
par(mfrow = c(1, 1))

#### model outputs ####

#print model output 
summary(pgls_edsvl)

#quick plot of pgls
plot(log10(eye_av) ~ log10(svl_av), data = av.edsvl)
abline(pgls_edsvl)

#Likelihood plot for Pagel's lambda from the PGLS model of eye diameter vs. the cuberoot of mass. Solid red line indicates estimate for lambda and broken red lines indcaite the 95% confidence interval
lambda.edsvl <- pgls.profile(pgls_edsvl, "lambda")
plot(lambda.edsvl)

```

For the PGLS of log mean eye diameter vs. log snout-vent length, Pagel's lambda was `r pgls.confint(pgls_edsvl, "lambda")$opt` with 95% confidence intervals of `r pgls.confint(pgls_edsvl, "lambda")$ci.val`, indicating that there is high phylogenetic signal in the residuals of our PGLS model fit (which is our measure of relative eye investment).

The diagnostic plots show that the data conformed to most model assumptions, though we do have two species with studentized residuals exceeding ±3, which could disproportionately influence the regression. To check whether this is happening, we re-ran the model with these two points eliminated. 