---
title: "Salamander eye scaling and investment"
author: "Nicole Bednarik & Katie Thomas"
date: 8 August 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(smatr)
library(picante)
library(nlme)
library(ape)
library(evobiR)
library(geiger)
library(caper)
library(phytools)
library(phylotools)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
#library(ggtree)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(mvMORPH)
#library(evomap)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

First we import our tidied salamander morphology & trait dataset. 

```{r load-data}

# Import tidied salamander eye size and trait data
salamanders <- data.frame(read.csv("../Data/Tidy/salamanders_tidy.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))

# Quick look at data structure
str(salamanders)
```

Next, we import the amphibian tree from Jetz and Pyron (2019) that has been pruned to match the salamander species in this dataset.

```{r, fig.height=18, fig.width=7}

#Import pruned phylogeny
caudatatree <- read.nexus("../Data/Tidy/caudata-tree.nex")

#Plot tree
plot.phylo(caudatatree, show.tip.label=TRUE, cex = 0.7)

```

# Interspecific allometry using PGLS

Here, we will use phylogenetic least-squares (PGLS) regressions to see how eyes are scaling with body size across different species of salamanders with different ecologies. First, we find species means for various morphological parameters.

```{r species-av}
#### species means for eye diameter and SVL ####
av.edsvl <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(SVL_mm))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), svl_av = mean(SVL_mm), n = n())

## Merge data means with other column info
av.edsvl <- merge(av.edsvl, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.edsvl <- av.edsvl %>% select(Genus_Species, eye_av, svl_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)

#### species means for eye diameter and cube root of mass ####
av.edmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(eyemean) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.edmass <- merge(av.edmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only, keeping only trait columns
av.edmass <- av.edmass %>% select(Genus_Species, eye_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)

#### species means for cube root of mass and SVL ####
av.svlmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(SVL_mm) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(svl_av = mean(SVL_mm), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.svlmass <- merge(av.svlmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.svlmass <- av.svlmass %>% select(Genus_Species, svl_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)
  
#### species means for cornea diameter and eye diameter ####
av.cded <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(eyemean))) %>% 
  group_by(Genus_Species) %>%
  summarise(cor_av = mean(cormean), eye_av = mean(eyemean), n = n())

## Merge data means with other column info
av.cded <- merge(av.cded, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.cded <- av.cded %>% select(Genus_Species, cor_av, eye_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)

#### species means for cornea diameter and SVL ####
av.cdsvl <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(SVL_mm))) %>% 
  group_by(Genus_Species) %>%
  summarise(cor_av = mean(cormean), svl_av = mean(SVL_mm), n = n())

## Merge data means with other column info
av.cdsvl <- merge(av.cdsvl, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.cdsvl <- av.cdsvl %>% select(Genus_Species, cor_av, svl_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)

#### species means for cornea diameter and cube root of mass ####
av.cdmass <- salamanders %>% 
  mutate_if(is.character, as.factor) %>% 
  filter(!(is.na(cormean) | is.na(rootmass))) %>% 
  group_by(Genus_Species) %>%
  summarise(cor_av = mean(cormean), rootmass_av = mean(rootmass), n = n())

## Merge data means with other column info
av.cdmass <- merge(av.cdmass, salamanders[match(unique(salamanders$Genus_Species), salamanders$Genus_Species), ], by="Genus_Species", all.x = TRUE, all.y = FALSE)

#Keep averages and associated traits only
av.cdmass <- av.cdmass %>% select(Genus_Species, cor_av, rootmass_av, n, Order, Suborder, Family, Subfamily, Genus, Species, Gill_Presence, Aquatic, Semiaquatic, Terrestrial, Scansorial, Fossorial, Subfossorial, Cave_dwelling, Nocturnal, Diurnal, Paedomorphic, Metamorphosizing, Direct_Development, Free_living, No_free_living, Lentic_water, Lotic_water, No_larva, Females_larger, Males_larger, No_SSD)



av.sal <- salamanders %>% 
  mutate_if(is.character, as.factor) %>%
  group_by(Genus_Species) %>%
  summarise(eye_av = mean(eyemean), 
            cor_av = mean(cormean),
            svl_av = mean(SVL_mm), 
            rootmass_av = mean(rootmass), 
            n = n())

```

In order to fit a PGLS regression in caper, we first need to make a comparative data object that includes our dataset and our phylogeny. Note that the gilled Pleurodeles waltl will be dropped from this, as there can only be one row for each species. 

```{r}

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.edsvl) <- av.edsvl$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.edsvl)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp1 <- comparative.data(phy = caudatatree, data = av.edsvl, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp1$dropped$tips #phylogeny
sal.comp1$dropped$unmatched.rows #dataset

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.edmass) <- av.edmass$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.edmass)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp2 <- comparative.data(phy = caudatatree, data = av.edmass, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp2$dropped$tips #phylogeny
sal.comp2$dropped$unmatched.rows #dataset

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.svlmass) <- av.svlmass$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.svlmass)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp3 <- comparative.data(phy = caudatatree, data = av.svlmass, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp3$dropped$tips #phylogeny
sal.comp3$dropped$unmatched.rows #dataset

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.cded) <- av.cded$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.cded)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp4 <- comparative.data(phy = caudatatree, data = av.cded, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp4$dropped$tips #phylogeny
sal.comp4$dropped$unmatched.rows #dataset

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.cdsvl) <- av.cdsvl$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.cdsvl)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp5 <- comparative.data(phy = caudatatree, data = av.cdsvl, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp5$dropped$tips #phylogeny
sal.comp5$dropped$unmatched.rows #dataset

#Make row names of dataset the species names (so it will match phylogeny tips)
rownames(av.cdmass) <- av.cdmass$Genus_Species

#check that names match in dataframe and tree
name.check(phy = caudatatree, data = av.cdmass)

#use caper function to combine phylogeny and data into one object (this function also matches species names in tree and dataset)
sal.comp6 <- comparative.data(phy = caudatatree, data = av.cdmass, 
                            names.col = Genus_Species, vcv = TRUE, 
                            na.omit = FALSE, warn.dropped = TRUE)

#check for dropped tips or dropped species
sal.comp6$dropped$tips #phylogeny
sal.comp6$dropped$unmatched.rows #dataset
```

# Relative eye size and eye investment in salamanders

To investigate how relative eye size differed across species with different ecological traits, we looked at the residuals from PGLS fits of cornea diameter vs. body size. 

First, we added residuals of PGLS fits to our datasets by species.


```{r pgls-residuals}
#PGLS cornea diameter vs. cube root of mass ------------
pgls_cdmass <- pgls(log10(cor_av) ~ log10(rootmass_av), 
               data = sal.comp6, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of cornea diameter ~ cube root of mass

#extract pgls residuals 
pglsres.cdmass <- residuals(pgls_cdmass) 

#name residuals
colnames(pglsres.cdmass) <- "pglsres.cdmass" 

#add rownames to original dataframe
rownames(av.cdmass) <- av.cdmass$mod_tiplabel 

#merge residuals with original data by rowname
av.cdmass.phy <- merge(av.cdmass, pglsres.cdmass, by = "row.names") 

#PGLS cornea diameter vs. SVL ------------
pgls_cdsvl <- pgls(log10(cor_av) ~ log10(svl_av), 
               data = sal.comp5, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#Residuals for PGLS of cornea diameter ~ SVL 

#extract pgls residuals
pglsres.cdsvl <- residuals(pgls_cdsvl) 

#name residuals
colnames(pglsres.cdsvl) <- "pglsres.cdsvl"

#add rownames to original dataframe
rownames(av.sal) <- av.sal$mod_tiplabel 

#merge residuals with original data by rowname
av.sal.phy <- merge(av.sal, pglsres.cdsvl, by = "row.names") 

#PGLS eye diameter vs. cornea diameter ------------
pgls_edcd <- pgls(log10(eye_av) ~ log10(cor_av), 
               data = sal.comp4, 
               lambda = "ML", #uses Maximum Likelihood estimate of lambda
               param.CI = 0.95)

#residuals for PGLS of eye diameter ~ cornea diameter 

#extract pgls residuals
pglsres.edcd <- residuals(pgls_edcd) 

#name residuals
colnames(pglsres.edcd) <- "pglsres.edcd" 

#add rownames to original dataframe
rownames(av.sal) <- av.sal$mod_tiplabel 

#merge residuals with original data by rowname
av.sal.phy <- merge(av.sal, pglsres.edcd, by = "row.names") 
```

## Phylogenetic distribution of absolute and relative eye size

We then plotted absolute eye size and relative eye investment (represented by the residual of the eye size vs. body size PGLS) onto the anuran phylogeny. 

```{r plot-phytools, fig.height = 15, fig.width = 8}

# color vectors -------

#create vector of colors for activity period
col_act <- c("Both" = "deeppink3",
             "Diurnal" = "darkgoldenrod1", 
             "Nocturnal" = "blueviolet",
             "Unknown" = "gray")

# create vector of colors for life history
col_lif <- c("Free-living larvae" = "darkturquoise", 
         "No free-living larvae" = "black", 
         "Unknown" = "gray")

# create vector of colors for larval habitat
col_larv <- c("Lentic water" = "royalblue1", 
         "Lotic water" = "skyblue1",
         "No larvae" = "lightcoral",
         "Obscured" = "darkslateblue",
         "On land" = "lightpink4",
         "Unknown" = "gray")

# create vector of colors for development
col_dich <- c("Direct Development" = "black", 
         "Metamorphosizing" = "chartreuse2",
         "Paedomorphic" = "gray")


# Prep data and phylogeny -----

#subset data for cornea diameter and mass
sal.bars <- av.sal.phy %>%
  mutate(oldtip = mod_tiplabel, 
         newtip = genus_species, 
         abseye = cor_av, 
         inveye = pglsres.cormass, 
         hab = Adult_habitat,
         act = Activity_period, 
         lif = Life_history,
         larv = Larval_habitat,
         dev = Development,
         family = Family) %>%
  select(oldtip, newtip, abseye, inveye, hab, act, mat, lif, larv, dev, family)

# set row names in dataset to match the tip labels in the tree
row.names(cormass.bars) <- cormass.bars$newtip

#drop phylogeny tips not in dataset)
tree_cormass <- drop.tip(frog.tree, frog.tree$tip.label[!(frog.tree$tip.label %in% cormass.bars$oldtip)])

#ladderize tree
tree_edmass <- ladderize(tree_cormass)

# Make Genus_species tip labels based on  Amphibian Species of the World names (and no family at end)
sp.tips <- data.frame(old = av.edmass$mod_tiplabel, new = av.edmass$genus_species)

# replace tip labels in phylogeny
tree_edmass <- sub.taxa.label(tree_edmass, sp.tips)

#check that phylogeny and data match exactly
name.check(tree_edmass, edmass.bars)

#resort trait dataset to the order of tree tip labels
edmass.bars <- edmass.bars[tree_edmass$tip.label, ] 

#make trait vector for absolute eye size
aveye <- as.vector(edmass.bars$abseye) 

#add tip label names to vector
names(aveye) <- edmass.bars$newtip 

#make trait vector for eye investment (PGLS residuals)
inv.mass <- as.vector(edmass.bars$inveye) #residuals of pgls
names(inv.mass) <- edmass.bars$newtip

#make trait vector of habitats
habs.edmass <- as.vector(edmass.bars$hab) 

#make vector of colors corresponding to phylogeny tips
tipcols.hab <- unname(col_hab[habs.edmass]) 


# Phylogeny with absolute eye diameters -----

#plot tree with eye size bars ant tip labels
plotTree.wBars(tree_edmass, aveye, 
               scale = 0.04, 
               tip.labels = TRUE, 
               offset = 0.35,
               ftype = "bi",
               fsize = 0.8,
               col = tipcols.hab)

#add legend for habitat states
legend(x = "left", legend = c("Aquatic", "Fossorial", "Ground-dwelling", "Scans orial", "Semiaquatic", "Subfossorial"), pch = 22, pt.cex= 2, pt.bg = col_hab, cex = 1, bty = "n", horiz = F)

# Phylogeny with relative eye size (compared to mass) -------

#plot tree with relative eye size bars
plotTree.wBars(tree_edmass, inv.mass, 
               scale = 0.7, 
               tip.labels = FALSE, 
               col = tipcols.hab,
               plot = FALSE, 
               add = TRUE)

```