---
title: "Salamander eye scaling and investment - boxplots"
author: "Nicole Bednarik & Katie Thomas"
date: 9 September 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(geiger)
library(caper)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

First we import our tidied salamander morphology & trait binary dataset with the residuals of a cornea ~ cube root of mass PGLS as a column.

```{r load-data}

# Import tidied salamander eye size and trait data
sal.res <- data.frame(read.csv("../Data/salamanders_residuals_binary.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
sal.res.2 <- data.frame(read.csv("../Data/salamanders_residuals_updated.csv", header=TRUE, na.strings=c("", "NA", " ", "<1")))
aquaticspecies <- data.frame(read.csv("../Data/aquaticspecies.csv", header=TRUE, na.strings=c("", "NA", " ", "<1")))

# Quick look at data structure
str(sal.res)
str(sal.res.2)

sal.res <- sal.res[-c(35:460)]

```

# Habitat plot

```{r}
#see what the levels of development are named
levels(as.factor(sal.res.2$Activity.Period))

#define a colorblind-friendly vector of colors for adult habitat
col_haba <- c("lentic water" = "chartreuse3",
             "lotic water" = "steelblue3")

#now you can see that each habitat is assigned a hex color when you look at the vector
col_haba

#make a vector of point shapes for adult habitat
shape_haba <- c("direct development" = 25,
             "metamorphosizing" = 24,
             "paedomorphic" = 23)

#see that each state has a shape now
shape_haba
```

```{r}
#boxplot of pgls residuals vs. aquatic species water environment (relative eye size)
habitat.rel <- ggplot(data = filter(aquaticspecies, !is.na(Adult.Habitat)), aes(x = reorder(Adult.Habitat, pglsres.cdmass, fun=mean), y=pglsres.cdmass, fill=Adult.Habitat, shape=Adult.Habitat)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha=0.5) +
  stat_summary(fun.y = mean, colour="black", geom="point", size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Species, shape = Adult.Habitat), size = 1.3, alpha = 0.3, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Relative cornea size") +
  xlab("Adult Habitat for Aquatic Species") +
  theme(legend.position = "none") +
# use our custom colors for habitat
  scale_fill_manual(values = col_haba, name = "Adult Habitat") +
# use our custom shapes for points
  scale_shape_manual(values = shape_haba, name = "Adult Habitat") 

#print plot
habitat.rel

#look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
library(plotly)
ggplotly(habitat.rel)

```

# Activity period plot

(I think you started editing this one for development but you have that farther down so maybe this one should be activity period? No colors defined for activity period)

```{r}
#boxplot of pgls residuals vs. activity period (relative eye size)
development.rel <- ggplot(data = filter(sal.res.2, !is.na(Activity.Period)), aes(x = reorder(Activity.Period, pglsres.cdmass, fun=mean), y=pglsres.cdmass, fill=Activity.Period, shape=Activity.Period)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha=0.5) +
  stat_summary(fun.y = mean, colour="black", geom="point", size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Development), size = 1.3, alpha = 0.3, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Relative cornea size") +
  xlab("Development") +
  theme(legend.position = "none") +
# use our custom colors for habitat
  scale_fill_manual(values = col_dev, name = "Development") +
# use our custom shapes for points
  scale_shape_manual(values = shape_dev, name = "Development") 

#print plot
development.rel

#look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
library(plotly)
ggplotly(development.rel)

```

# Development plot 

```{r}
#see what the levels of development are named
levels(as.factor(sal.res.2$Development))

#define a colorblind-friendly vector of colors for adult habitat
col_dev <- c("direct development" = "chartreuse3",
             "metamorphosizing" = "steelblue3",
             "paedomorphic" = "slateblue1")

#now you can see that each habitat is assigned a hex color when you look at the vector
col_dev

#make a vector of point shapes for adult habitat
shape_dev <- c("direct development" = 25,
             "metamorphosizing" = 24,
             "paedomorphic" = 23)

#see that each state has a shape now
shape_dev
```

```{r}
#boxplot of pgls residuals vs. development (relative eye size)
development.rel <- ggplot(data = filter(sal.res.2, !is.na(Development)), aes(x = reorder(Development, pglsres.cdmass, fun=mean), y=pglsres.cdmass, fill=Development, shape=Development)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha=0.5) +
  stat_summary(fun.y = mean, colour="black", geom="point", size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Development), size = 1.3, alpha = 0.3, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Relative cornea size") +
  xlab("Development") +
  theme(legend.position = "none") +
# use our custom colors for habitat
  scale_fill_manual(values = col_dev, name = "Development") +
# use our custom shapes for points
  scale_shape_manual(values = shape_dev, name = "Development") 

#print plot
development.rel

#look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
library(plotly)
ggplotly(development.rel)

```

# Color  figures with pallettes in the frog paper

Here I will just re-run the code to make the figures you want for your poster, with a custom color pallette so that the habitat categories will be the same for all of them. 

First we define what we want that color pallette to be:

```{r}
#see what the levels of adult habitat are named
levels(as.factor(sal.res.2$Adult.Habitat))

#define a colorblind-friendly vector of colors for adult habitat
col_hab <- c("aquatic" = "#0072B2",
             "scansorial" = "#009E73",
             "semiaquatic" = "#56B4E9",
             "subfossorial" = "#CC79A7",
             "terrestrial" = "#E69F00")

#now you can see that each habitat is assigned a hex color when you look at the vector
col_hab

#make a vector of point shapes for adult habitat
shape_hab <- c("aquatic" = 25,
             "scansorial" = 24,
             "semiaquatic" = 22,
             "subfossorial" = 21,
             "terrestrial" = 23)

#see that each state has a shape now
shape_hab
```

## Boxplot adult habitat figure

You've already made a nice boxplot of adult habitats and relative eye size, so now we can just modify that plot to make it nicer, ordered sensibly, and colored how we want. 

```{r}

#boxplot of pgls residuals vs. adult habitat (relative eye size)
adult_habitat.rel <- ggplot(data = filter(sal.res.2, !is.na(Adult.Habitat)), aes(x = reorder(Adult.Habitat, pglsres.cdmass, fun = mean), y = pglsres.cdmass, fill=Adult.Habitat, shape = Adult.Habitat)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, colour="black", geom="point", size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Adult.Habitat, color=Development),  size = 2, alpha = 0.8, position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Relative cornea size") +
  xlab("Adult habitat") +
  theme(legend.position = "none") +
# use our custom colors for habitat
  scale_fill_manual(values = col_hab, name = "Adult habitat") +
# use our custom shapes for points
  scale_shape_manual(values = shape_hab, name = "Adult habitat") 
  

#print plot
adult_habitat.rel

#look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
library(plotly)
ggplotly(adult_habitat.rel)

#export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
pdf(file = "boxplots.pdf", height=4, width=6)
adult_habitat.rel
dev.off()

```
