---
title: "Salamander eye scaling and investment - boxplots"
author: "Nicole Bednarik & Katie Thomas"
date: 9 September 2022
output:
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_fold: show
---

```{r setup, include = FALSE}

# Load package libraries
library(plyr)
library(geiger)
library(caper)
library(plotly)
library(kableExtra)
library(gridExtra)
library(ggimage)
library(grid)
library(gtable)
library(cowplot)
library(knitr)
library(tidyverse)

# Markdown settings
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.height = 4.5, fig.width=8)

```

# Data

First we import our tidied salamander morphology & trait binary dataset with the residuals of a cornea ~ cube root of mass PGLS as a column.

```{r load-data}

# Import tidied salamander eye size and trait data
sal.res <- data.frame(read.csv("../Data/salamanders_residuals_binary.csv", header=TRUE, na.strings=c("", "NA", " ","<1")))
sal.res.2 <- data.frame(read.csv("../Data/salamanders_residuals_updated.csv", header=TRUE, na.strings=c("", "NA", " ", "<1")))

# Quick look at data structure
str(sal.res)
str(sal.res.2)

sal.res <- sal.res[-c(35:460)]

```

# Color  figures with pallettes in the frog paper

Here I will just re-run the code to make the figures you want for your poster, with a custom color pallette so that the habitat categories will be the same for all of them. 

First we define what we want that color pallette to be:

```{r}
#see what the levels of adult habitat are named
levels(as.factor(sal.res.2$Adult.Habitat))

#define a colorblind-friendly vector of colors for adult habitat
col_hab <- c("aquatic" = "#0072B2",
             "scansorial" = "#009E73",
             "semiaquatic" = "#56B4E9",
             "subfossorial" = "#CC79A7",
             "terrestrial" = "#E69F00")

#now you can see that each habitat is assigned a hex color when you look at the vector
col_hab

#make a vector of point shapes for adult habitat
shape_hab <- c("aquatic" = 25,
             "scansorial" = 24,
             "semiaquatic" = 22,
             "subfossorial" = 21,
             "terrestrial" = 23)

#see that each state has a shape now
shape_hab
```

## Boxplot adult habitat figure

You've already made a nice boxplot of adult habitats and relative eye size, so now we can just modify that plot to make it nicer, ordered sensibly, and colored how we want. 

```{r}

#boxplot of pgls residuals vs. adult habitat (relative eye size)
adult_habitat.rel <- ggplot(data = filter(sal.res.2, !is.na(Adult.Habitat)), aes(x = reorder(Adult.Habitat, pglsres.cdmass, fun = mean), y = pglsres.cdmass, fill=Adult.Habitat, shape = Adult.Habitat)) +
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.5) +
  stat_summary(fun.y = mean, colour="black", geom="point", size=3, show_guide = FALSE) +
  geom_jitter(aes(text = Genus_Species, shape = Adult.Habitat),  size = 1.3, alpha = 0.3, col = "black", fill = "black", position = position_jitter(0.15)) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black")) +
  theme(axis.text = element_text(angle = 0, size = 9)) +
  ylab("Relative cornea size") +
  xlab("Adult habitat") +
  theme(legend.position = "none") +
# use our custom colors for habitat
  scale_fill_manual(values = col_hab, name = "Adult habitat") +
# use our custom shapes for points
  scale_shape_manual(values = shape_hab, name = "Adult habitat")

#print plot
adult_habitat.rel

#look at the plot with the plotly package - you can hover over data points to see what each species is, great way to check outliers!
library(plotly)
ggplotly(adult_habitat.rel)

#export plot to pdf - you can change the export dimensions by changing the height and width numbers to make it fit well on your poster!
pdf(file = "boxplots.pdf", height=4, width=6)
adult_habitat.rel
dev.off()

```